### 2015-07-06 poisson_errker.tex
Correcting some mistakes

\subsection{Approximate Solution}
We are interested in the scenario where the function $f$ is only known
through its point samples. Specifically, we assume that the samples of
$f$ reside on the sites of the lattice $\lattice{L}_h$. For the
purpose of enumerating the samples that are contained within
$\cube^s$, let us define the point set
\begin{equation}
  \pointset{T}_h := \underbrace{\{\vect{x}_1,\ldots,\vect{x}_{2^sN} \}}_{\pointset{I}_h}
  \cup
\underbrace{\{\vect{x}_{2^sN + 1},\ldots,\vect{x}_{2^sN + M} \}}_{\pointset{B}_h},
\end{equation}
where $\pointset{I}_h := \ocube^s \cap \{\vect{x} + \vect{m} :
\vect{x} \in \lattice{L}_h \cap \oucube^s, \vect{m} \in \ints^s \}$
consists of interior lattice points, while $\pointset{B}_h$ consists
of extended boundary points, i.e. $\pointset{B}_h := (\lattice{L}_h
\cap \partial\cube^s \cap (-1,1]^s) \cup (\lattice{L}_h \cap
\ocube^s\backslash \pointset{I}_h)$. Two dimensional illustrations are
shown in Figure~\ref{fig:pointset}. We note that with the above
definition of $\pointset{I}_h$, the points $\vect{x}_j$ for $j \in
\{1, \ldots, N\}$ are contained in the open unit cube $\oucube^s$
whereas the remaining points ($\vect{x}_j$ for $j \in \{N+1, \ldots,
2^sN\}$) lie outside. Furthermore, we assume
that the lattice $\lattice{L}$ and the sampling rate $h$ are such that
the set $\{\vect{x} + 2\vect{m} : \vect{x} \in \pointset{T}_h,
\vect{m} \in \ints^s\} = \lattice{L}_h$. We remark that this
requirement for $\lattice{L}_h$ is also
satisfied by integration lattices that are commonly used
to devise quadrature rules within the unit cube $\ucube^s$~\cite{nied92}.
\begin{figure}[t]
  \centering
  \subfloat[Cartesian]{\label{fig:cartesian}\input{./figures/2DCartesian}}
  \hspace{2em}
  \subfloat[Quincunx]{\label{fig:quincunx}\input{./figures/2DQuincunx}}
  \caption[Interior and boundary samples]{Illustration of the point set
  $\pointset{T}_h$ on the two
    dimensional Cartesian ($\mat{L} = \bigl( \begin{smallmatrix} 1&0\\
      0&1
    \end{smallmatrix} \bigr)$, $h=\tfrac{1}{8}$) and Quincunx ($\mat{L} = \bigl(
    \begin{smallmatrix} 1&-1\\
      1&1
    \end{smallmatrix} \bigr)$, $h=\tfrac{1}{10}$) lattices. The interior points
    ($\bullet$) belong to the set $\pointset{I}_h$ while the extended
    boundary points ($\circ$) belong to $\pointset{B}_h$.}
  \label{fig:pointset}
\end{figure}

@@

We wish to use the samples of $f$ to approximate the function $V$ that solves the Poisson equation~\eqref{eq:PoissonEquation}. Since
$\chi$ is a periodic function, we follow the recipe of Jacob \textit{et al.}~\cite{jacob02} and seek an approximation that lies in a space generated by a periodic reconstruction function. Specifically, we are interested in the case
where $\appf{\chi}$ lies in the space $\aspace{\lattice{L}_h}{\varphi_p} := \mathrm{span}_{\vect{x}_j\in\pointset{T}_h}\{
\varphi_p(\frac{\vect{x}-\vect{x}_j}{h}) \}$ that is spanned by the scaled and translated versions of a periodic function $\varphi_p$. Our approximation is given by
\begin{equation}
  \appf{\chi}(\vect{x}) = \sum_{\vect{x}_j \in \pointset{T}_h} c[\vect{x}_j] 
\varphi_p\bigl(\frac{\vect{x} - \vect{x}_j}{h}\bigr),
\label{eq:ansatz}
\end{equation}
where $c[\vect{x}_j]$ is an unknown coefficient sequence defined on the point set $\pointset{T}_h$ that is to be determined from the samples of $f$. The function $\varphi_p$ is a periodized version of a generating function $\varphi$ and is defined as
\begin{equation}
\varphi_p(\vect{x}) := \sum_{\vect{m} \in \ints^s} 
\varphi\bigl(\vect{x} - \frac{2}{h} \vect{m} \bigr).
\label{eq:periodization}
\end{equation}

@@

In many approximation scenarios, one typically assumes a Dirac point sampling
model, i.e. $\analysis{\varphi}(\vect{x}) = \delta(\vect{x})$.
This prevents the direct realization of the minimum approximation scenario.
Furthermore, for many functions encountered in practice, the power spectrum is
concentrated around $\vect{\omega} = 0$. For such scenarios, one speaks of an
\emph{asymptotically optimal} $k$-th order approximation scheme if the $L_2$
error behaves as $O(h^k)$ where $k = \order{\lattice{L}}{\varphi}$. Our operator
discretization approach is also based on these assumptions. Provided that $V \in
W_2^k(\cube^s)$, our goal is to design suitable digital filters that yield
$\appf{V}$ such that $\LLnorm{V - \appf{V}} = O(h^k)$. As discussed earlier
in~\SC{sec:interpQuasiInterp}, the criterion that needs to be satisfied is
$E_{\mathsf{mod}}(\vect{\omega}) = O(\norm{\vect{\omega}}^{2k})$.

\subsection{Relationship to the Galerkin method}
\label{sec:Galerkin}
The modified residue error kernel~\eqref{eq:opResidue} can also be analyzed in
light of the Galerkin method~\cite{quarteroni08} using a weak formulation of the Poisson
equation~\eqref{eq:PoissonEquation}, where the trial and test spaces are the
same with the notable difference that the spaces are not required to explicitly
satisfy any particular boundary conditions. Rather, a zero boundary condition is
implicitly obtained by requiring that the solution coefficients $c[\cdot]$ be
odd as explained in~\SC{sec:approximateSolution}. Here, we establish the
connection for the homogeneous case but the analysis also easily extends to the
non-homogeneous case, as well as to other types of differential operators.

In its weak form, the solution $V \in L_2(\cube^s)$ to the homogeneous
Poisson equation~\eqref{eq:PoissonEquation} satisfies the weak
formulation
\begin{equation}
  \innerproduct{V}{u} = \mathcal{F}(u), \;\; \forall u \in L_2(\cube^s),
\end{equation}
where the functions $u$ are suitable test functions. The functional
$\mathcal{F}(\cdot)$ is defined as $\mathcal{F}(u) :=
\innerproduct{\invlaplacian f}{u}$, where $f$ is the odd extension of
the function that appears on the right hand side
of~\eqref{eq:PoissonEquation}, and the operator
$\stpairs{\invlaplacian}{\tfrac{1}{-\pi^2\norm{\vect{m}}^2}}$ has the
interpretation of the inverse Laplacian as introduced
in~\eqref{eq:fourierSolution}. We remark that this formulation is
slightly stronger than the usual weak formulations based on the
Laplacian $\laplacian$ where the trial and test spaces coincide with
the Sobolev space $W_2^1(\cube^s)$. In comparison, here the trial and
test spaces coincide with the more general space $L_2(\cube^s)$. This
formulation is also in direct correspondence with our earlier
treatment of the problem.

Let us now restrict attention to the finite dimensional space
$\aspace{\lattice{L}_h}{\varphi_p} \subset L_2(\cube^s)$. We seek a
weak solution $\appf{V} \in \aspace{\lattice{L}_h}{\varphi_p}$ so that
\begin{equation}
  \innerproduct{\appf{V}}{u_h} = \mathcal{F}(u_h), \;\;
  \forall u_h \in \aspace{\lattice{L}_h}{\varphi_p},
\end{equation}
and the Galerkin residual satifies the orthogonality condition
\begin{equation}
  \innerproduct{V - \appf{V}}{u_h} = 0, \;\;
  \forall u_h \in \aspace{\lattice{L}_h}{\varphi_p}.
  \label{eq:GalerkinOrthogonality}
\end{equation}
From this, it can be deduced that the minimum error approximation
$\appf{V}$ is the orthogonal projection of $V$ upon
$\aspace{\lattice{L}_h}{\varphi_p}$. The coefficients of the
approximation are thus given by
\begin{equation}
  \begin{split}
    c[\vect{x}_j] & = \mathcal{F}(h^{-s}
    \dual{\varphi}(\frac{\vect{x}-\vect{x}_j}{h})) \\
    & = 
    \innerproduct{f(\vect{x})}{\invlaplacian(h^{-s}\dual{\varphi}(\frac{\vect{x}-\vect{x}_j}{h}))}.
  \end{split}
\end{equation}

Observe that if we use the analysis function $\analysis{\varphi}_p =
\invlaplacian \dual{\varphi}_p$ (in a distributional sense) in the
measurement equation~\eqref{eq:measurements}, and subsequently employ
no filtering (i.e. $\gamma[\vect{x}_j] = \delta[\vect{x}_j]$
in~\eqref{eq:opApproximation}), then the measurements obtained will realize the orthogonal projection of $V$
upon $\aspace{\lattice{L}_h}{\varphi_p}$. In this case, the modified Fourier
residue error kernel~\eqref{eq:opResidue} vanishes,
i.e. $E_{\mathsf{mod}}(\vect{\omega}) = 0$.

>>>>

\subsection{Approximate Solution} \sec{sub:approximatesol}
We are interested in the scenario where the function $f$ is only known
through its point samples. Specifically, we assume that the samples of
$f$ reside on the sites of the lattice $\lattice{L}_h$. For the
purpose of enumerating the samples that are contained within
$\cube^s$, let us define the point set
\begin{equation}
  \pointset{T}_h := \underbrace{\{\vect{x}_1,\ldots,\vect{x}_{2^sN} \}}_{\pointset{I}_h}
  \cup
\underbrace{\{\vect{x}_{2^sN + 1},\ldots,\vect{x}_{2^sN + M} \}}_{\pointset{B}_h},
\end{equation}
where $\pointset{I}_h := \ocube^s \cap \{\vect{x} + \vect{m} :
\vect{x} \in \lattice{L}_h \cap \oucube^s, \vect{m} \in \ints^s \}$
consists of interior lattice points, while $\pointset{B}_h$ consists
of extended boundary points, i.e. $\pointset{B}_h := (\lattice{L}_h
\cap \partial\cube^s \cap (-1,1]^s) \cup (\lattice{L}_h \cap
\ocube^s\backslash \pointset{I}_h)$. Two dimensional illustrations are
shown in Figure~\ref{fig:pointset}. We note that with the above
definition of $\pointset{I}_h$, the points $\vect{x}_j$ for $j \in
\{1, \ldots, N\}$ are contained in the open unit cube $\oucube^s$
whereas the remaining points ($\vect{x}_j$ for $j \in \{N+1, \ldots,
2^sN\}$) lie outside. Furthermore, we assume
that the lattice $\lattice{L}$ and the sampling rate $h$ are such that
the set $\{\vect{x} + 2\vect{m} : \vect{x} \in \pointset{T}_h,
\vect{m} \in \ints^s\} = \lattice{L}_h$. We remark that this
requirement for $\lattice{L}_h$ is also
satisfied by integration lattices that are commonly used
to devise quadrature rules within the unit cube $\ucube^s$~\cite{nied92}.
\begin{figure}[t]
  \centering
  \subfloat[Cartesian]{\label{fig:cartesian}\input{./figures/2DCartesian}}
  \hspace{2em}
  \subfloat[Quincunx]{\label{fig:quincunx}\input{./figures/2DQuincunx}}
  \caption[Interior and boundary samples]{Illustration of the point set
  $\pointset{T}_h$ on the two
    dimensional Cartesian ($\mat{L} = \bigl( \begin{smallmatrix} 1&0\\
      0&1
    \end{smallmatrix} \bigr)$, $h=\tfrac{1}{8}$) and Quincunx ($\mat{L} = \bigl(
    \begin{smallmatrix} 1&-1\\
      1&1
    \end{smallmatrix} \bigr)$, $h=\tfrac{1}{10}$) lattices. The interior points
    ($\bullet$) belong to the set $\pointset{I}_h$ while the extended
    boundary points ($\circ$) belong to $\pointset{B}_h$.}
  \label{fig:pointset}
\end{figure}

@@

We wish to use the samples of $f$ to approximate the function $\chi$ that solves the Poisson equation~\eqref{eq:PoissonEquation}. Since
$\chi$ is a periodic function, we follow the recipe of Jacob \textit{et al.}~\cite{jacob02} and seek an approximation that lies in a space generated by a periodic reconstruction function. Specifically, we are interested in the case
where $\appf{\chi}$ lies in the space $\aspace{\lattice{L}_h}{\varphi_p} := \mathrm{span}_{\vect{x}_j\in\pointset{T}_h}\{
\varphi_p(\frac{\vect{x}-\vect{x}_j}{h}) \}$ that is spanned by the scaled and translated versions of a periodic function $\varphi_p$. Our approximation is given by
\begin{equation}
  \appf{\chi}(\vect{x}) = \sum_{\vect{x}_j \in \pointset{T}_h} c[\vect{x}_j] 
\varphi_p\bigl(\frac{\vect{x} - \vect{x}_j}{h}\bigr),
\label{eq:ansatz}
\end{equation}
where $c[\vect{x}_j]$ is an unknown coefficient sequence defined on the point set $\pointset{T}_h$ that is to be determined from the samples of $f$. The function $\varphi_p$ is a periodized version of a generating function $\varphi$ and is defined as
\begin{equation}
\varphi_p(\vect{x}) := \sum_{\vect{m} \in \ints^s} 
\varphi\bigl(\vect{x} - \frac{2}{h} \vect{m} \bigr).
\label{eq:periodization}
\end{equation}

@@

In many approximation scenarios, one typically assumes a Dirac point sampling
model, i.e. $\analysis{\varphi}(\vect{x}) = \delta(\vect{x})$.
This prevents the direct realization of the minimum approximation scenario.
Furthermore, for many functions encountered in practice, the power spectrum is
concentrated around $\vect{\omega} = 0$. For such scenarios, one speaks of an
\emph{asymptotically optimal} $k$-th order approximation scheme if the $L_2$
error behaves as $O(h^k)$ where $k = \order{\lattice{L}}{\varphi}$. Our operator
discretization approach is also based on these assumptions. Provided that $\chi$ has appropriate smoothness, our goal is to design suitable digital filters that yield %$V \in W_2^k(\cube^s)$, our goal is to design suitable digital filters that yield
$\appf{\chi}$ such that $\LLnorm{\chi - \appf{chi}} = O(h^k)$. %As discussed earlier in~\SC{sec:interpQuasiInterp}, the criterion that needs to be satisfied is $E_{\mathsf{mod}}(\vect{\omega}) = O(\norm{\vect{\omega}}^{2k})$.

\subsection{Relationship to the Galerkin method}
\label{sec:Galerkin}
The modified residue error kernel~\eqref{eq:opResidue} can also be analyzed in light of the Galerkin method~\cite{quarteroni08} using a weak formulation of the Poisson equation~\eqref{eq:PoissonEquation}, where the trial and test spaces are the same with the notable difference that the spaces are not required to explicitly satisfy any particular boundary conditions. Rather, a zero boundary condition is implicitly obtained by requiring that the solution coefficients $c[\cdot]$ be odd as explained in~\SC{sub:approximatesol}. Here, we establish the connection for the homogeneous case but the analysis also easily extends to the non-homogeneous case, as well as to other types of differential operators.

In its weak form, the solution $\chi \in L_2(\cube^s)$ to the homogeneous Poisson equation~\eqref{eq:PoissonEquation} satisfies the weak formulation
\begin{equation}
  \innerproduct{\chi}{u} = \mathcal{F}(u), \;\; \forall u \in L_2(\cube^s),
\end{equation}
where the functions $u$ are suitable test functions. The functional
$\mathcal{F}(\cdot)$ is defined as $\mathcal{F}(u) :=
\innerproduct{\invlaplacian f}{u}$, where $f$ is the odd extension of
the function that appears on the right hand side
of~\eqref{eq:PoissonEquation}, and the operator
$\stpairs{\invlaplacian}{\tfrac{1}{-\pi^2\norm{\vect{m}}^2}}$ has the
interpretation of the inverse Laplacian as introduced
in~\eqref{eq:fourierSolution}. We remark that this formulation is
slightly stronger than the usual weak formulations based on the
Laplacian $\laplacian$ where the trial and test spaces coincide with
the Sobolev space $W_2^1(\cube^s)$. In comparison, here the trial and
test spaces coincide with the more general space $L_2(\cube^s)$. This
formulation is also in direct correspondence with our earlier
treatment of the problem.

Let us now restrict attention to the finite dimensional space
$\aspace{\lattice{L}_h}{\varphi_p} \subset L_2(\cube^s)$. We seek a
weak solution $\appf{\chi} \in \aspace{\lattice{L}_h}{\varphi_p}$ so that
\begin{equation}
  \innerproduct{\appf{\chi}}{u_h} = \mathcal{F}(u_h), \;\;
  \forall u_h \in \aspace{\lattice{L}_h}{\varphi_p},
\end{equation}
and the Galerkin residual satifies the orthogonality condition
\begin{equation}
  \innerproduct{\chi - \appf{\chi}}{u_h} = 0, \;\;
  \forall u_h \in \aspace{\lattice{L}_h}{\varphi_p}.
  \label{eq:GalerkinOrthogonality}
\end{equation}
From this, it can be deduced that the minimum error approximation $\appf{\chi}$ is the orthogonal projection of $\chi$ upon $\aspace{\lattice{L}_h}{\varphi_p}$. The coefficients of the approximation are thus given by
\begin{equation}
  \begin{split}
    c[\vect{x}_j] & = \mathcal{F}(h^{-s}
    \dual{\varphi}(\frac{\vect{x}-\vect{x}_j}{h})) \\
    & = 
    \innerproduct{f(\vect{x})}{\invlaplacian(h^{-s}\dual{\varphi}(\frac{\vect{x}-\vect{x}_j}{h}))}.
  \end{split}
\end{equation}

Observe that if we use the analysis function $\analysis{\varphi}_p = \invlaplacian \dual{\varphi}_p$ (in a distributional sense) in the measurement equation~\eqref{eq:measurements}, and subsequently employ no filtering (i.e. $\gamma[\vect{x}_j] = \delta[\vect{x}_j]$ in~\eqref{eq:opApproximation}), then the measurements obtained will realize the orthogonal projection of $\chi$ upon $\aspace{\lattice{L}_h}{\varphi_p}$. In this case, the modified Fourier residue error kernel~\eqref{eq:opResidue} vanishes, i.e. $E_{\mathsf{mod}}(\vect{\omega}) = 0$.
